AMOVA
===


Load in data and read STRUCTURE infiles
```{r, include=FALSE}
library(adegenet) #read.structure & basic stats
library(hierfstat) #pairwise Fst
library(ggplot2) #plotting
library(poppr)
```

```{r}
allr<-read.structure("allcra473_240_in.stru", n.ind=473, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=85, ask=FALSE, quiet=FALSE)
all<-read.structure("all480_240_in.stru", n.ind=480, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=87, ask=FALSE, quiet=FALSE)
eu<-read.structure("eu199_240_in.stru", n.ind=199, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1, NA.char="-9", pop=36, ask=FALSE, quiet=FALSE)
nar<-read.structure("nacra232_240_in.stru", n.ind=232, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=42, ask=FALSE, quiet=FALSE)
au<-read.structure("au42_240_in.stru", n.ind=42, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=7,  ask=FALSE, quiet=FALSE)
```

Add data to genind object
```{r}
allother<-read.table("all476_other.txt",header=T)
all@other <- as.data.frame(allother[,1:6])
all@pop<-all@other[,1]
strata(all)<-data.frame(all@other[,1:4])
all@other$latlong<-data.frame(all@other[,5:6])

allrother<-read.table("allr470_other.txt",header=T)
allr@other <- as.data.frame(allrother[,1:6])
allr@pop<-allr@other[,1]
strata(allr)<-data.frame(allr@other[,1:4])
allr@other$latlong<-data.frame(allr@other[,5:6])

euother<-read.table("eu197_other.txt",header=T)
eu@other <- as.data.frame(euother[,1:6])
strata(eu)<-data.frame(eu@other[,1:4])
eu@pop<-eu@other[,1]
eu@other$latlong<-data.frame(eu@other[,5:6])

narother<-read.table("nar232_other.txt",header=T)
nar@other <- as.data.frame(narother[,1:6])
strata(nar)<-data.frame(nar@other[,1:4])
nar@pop<-nar@other[,1]
nar@other$latlong<-data.frame(nar@other[,5:6])

auother<-read.table("au41_other.txt",header=T)
au@other <- as.data.frame(auother[,1:6])
strata(au)<-data.frame(au$other[,1:4])
au@pop<-au@other[,1]
au@other$latlong<-data.frame(au@other[,5:6])
```

AMOVA shows at what level variation is explained. Here, I can look at variation within individuals; between individuals within populations, between populations within a range, and between ranges. 
Next, I will do an AMOVA without calculating within-sample variance. This within-sample variance is caused by heterozygosities within diploid individuals, the analysis splits the diploid genotypes into haplotypes and uses this to calculate variance. I am wondering though if biases in the data might skew this variance extremely.

Note that the "sample" in the output is actually the lowest level defined in the formula, in this case "pop".

{poppr,ade4,pegas}
http://grunwaldlab.github.io/Population_Genetics_in_R/AMOVA.html
```{r}
allr_gc <- as.genclone(allr)
allr_gc
table(strata(allr_gc, ~range/pop)) 
allr_amova<-poppr.amova(allr_gc, ~range/pop,missing="geno",within=FALSE,cutoff=.9)
```

Significance testing of this result
```{r}
set.seed(1999)
allr_amova_s   <- randtest(allr_amova, nrepet = 999)
```



Additionally, pegas is better at calculating the amova, but can only do this when within-sample variance is NOT calculated
```{r}
allr_amova_p<-poppr.amova(allr_gc, ~range/pop,method="pegas", within=FALSE, 
                          missing="geno", cutoff=.9)
allr_amova_sp   <- randtest(allr_amova_p, nrepet = 999) # doesn't work
```




