Outlier Analysis
===

> Code by Lotte van Boheemen

The below code will do the following:
- Look at basic distribution of the data  
- Define what an outlier is  
- Calculate number of outliers based on sliding window, where window size is set to a certain number of bases
- Print out actual window size in bases (between first and last SNP in window)
- Print out window size in SNPs (important if window length is limiting factor)
- Print out number of outliers in SNPs
- Plot distribution of outliers and actual window sizes within each provided data set (ranges)

> Update 25-6-16: Code is streamlined to run as a function, plot most important graphs and output results, so that input data doesn't have to be changed inside the code


```{r,include=FALSE}
library("knitr")
library("dplyr")
library("plyr")
library("dplyr")

opts_knit$set(root.dir = '/Users/lottevanboheemen/Documents/Monash/PhD/Analyses/Data')
```

---

**Read data**

Refgenome_data.txt is a file containing all the contigs of the reference genome, so also regions in which no SNPs are called
```{r}
cdata <- read.table("refgenome_data.txt",header=T)

#xtx_na contains xtx values averaged over 3 independent (different root) Bayenv2 runs using .5M iterations, for North-American sampling locations only
xdat_na <- read.table("xtx_na",header=T)
xdat_eu <- read.table("xtx_eu.txt",header=T)
xdat_au <- read.table("xtx_au.txt",header=T)
```


Add contig length information to xtx data by combining matrices {dplyr}
```{r}
x_na <- left_join(xdat_na,cdata,by="contig")
x_na$contig <- as.factor(x_na$contig)
x_na$myrank <- rank(x_na$xtx)/length(x_na$xtx)
x_na_100thperc <- (subset(x_na,myrank>=0.99))
write.table(subset(x_na,myrank>=0.99),"x_na_100thperc")
plot(x_na$xtx ~ x_na$myrank)
pdf("_Graph/xna-snps.pdf", height=6, width=6)
plot(x_na$xtx, col=ifelse(x_na$myrank>=0.99,"red","black"), ylim=c(30,80)) 
dev.off()
abline(v=.99)

x_eu <- left_join(xdat_eu, cdata,by="contig")
x_eu$contig<-as.factor(x_eu$contig)
x_eu$myrank <- rank(x_eu$xtx)/length(x_eu$xtx)
x_eu_100thperc <- (subset(x_eu,myrank>=0.99))
write.table(subset(x_eu,myrank>=0.99),"x_eu_100thperc")
plot(x_eu$xtx ~ x_eu$myrank)
pdf("_Graph/xeu-snps.pdf", height=6, width=6)
plot(x_eu$xtx, col=ifelse(x_eu$myrank>=0.99,"red","black"), ylim=c(30,80)) 
dev.off()
abline(v=.99)

x_au <- left_join(xdat_au, cdata,by="contig")
x_au$contig <- as.factor(x_au$contig)
x_au$myrank <- rank(x_au$xtx)/length(x_au$xtx)
x_au_100thperc <- (subset(x_au,myrank>=0.99))
write.table(x_au_100thperc,"x_au_100thperc")
plot(x_au$xtx ~ x_au$myrank)
abline(v=.99)
```

Calculate  stats
```{r}
#average length of each contig
summary(cdata$length)

# Calculate xtx outliers
summary(x_na$xtx)
summary(x_eu$xtx)
summary(x_au$xtx)
x11()
plot(x_au$xtx)
ggplot(x_au, aes(x= myrank, y=xtx)) + geom_point() +
geom_errorbar(aes(ymin=xtx-se, ymax=xtx+se), width=.05)

str(x_au)
```

---  

Outlier window where slide is based on position on the contig (so not SNP).

```{r}
#window length in bases
win.length <- 1000
# slide in bases
slide <- 1000
# define outlier value
#top <- 0.99

outlier.window <- function(data, output, top){
  
  top <- top
  ##Define what an outlier is, based on 99th percentile:
  
  par(mfrow=c(2,2))
  
  #show xtx frequency distribution
  hist(data$xtx)
  
  data$Klength <- data$length/1000
  
  # plot xtx vs length contig
  plot(data$xtx ~ data$Klength,xlab="Scaffold length (Kb)", ylab = "SNP XtX values")
  
  #identify the top percentile by ranking all the data according to their percentile ranks
  data$myrank<-rank(data$xtx)/length(data$xtx)
  
  #Check the distribution of percentiles vs xtx values
  plot(data$xtx~data$myrank, xlab = "Ranked XtX values", ylab = "XtX values")
  
  
  ## run the main part of the outlier window function:
  
  # create a vector including names of each contig for which SNPs have been found
  cont <- levels(factor(data$contig)) 
  
  # prepare the space for results to be written to
  results <- data.frame(SNPs=numeric(),contig=numeric(),win.start=numeric(),loc.start = numeric(), win.end = numeric(), index = numeric(), num.SNPs = numeric(), win.mean = numeric(), current.window = numeric(), num.outlier = numeric())
  
  SNPinfo <- data.frame(contig=numeric(),location = numeric(), xtx = numeric(), myrank = numeric (), win.start = numeric(), win.end = numeric())
  
  #set index for loops, this is the index of SNP within each contig
  i <- 1
  #set index for results, this is the index of window passed threshold
  j <- 1
  #set index for contigs
  k <- 1
  # set index for number of SNPs within each window (incrementing until set window length). l is a variable increasing in size to fit the maximum number it can reach depending on window length. i+l is the index of the last SNP in the window. l + 1 is the window size in number of SNP. 
  l <- 0
  #set index for SNPinfo results, where every line is a SNP and all information
  a <- 1
  
  # Loop through each contig, stop the loop at the final contig
  while(k <= length(cont)){
    
    # vector with xtx values for all SNPs in each contig
    x <- data$xtx[data$contig==cont[k]]
    
    # vector with location information for all SNPs in each contig
    y <- data$location[data$contig==cont[k]] 
    
    # vector with percentile ranks for each SNP based on xtx values
    z <- data$myrank[data$contig==cont[k]]
  
    # set start at first location within contig
    win.start <- y[i]
    
    # number of SNPs in contig
    snps <- length(x)
  
    #  Loop through all values within a contig, don't go past last value within the contig (all restrictions below are different ways of saying the same. However, the function doesn't work without all restrictions)
    while(i <= snps && (win.start + win.length) <= y[snps] && y[i] < (win.start+win.length)) {
      
      # set end position
      win.end <- win.start + win.length
      
      
      # add SNPs to the window until the temporary window is the maximum threshold it could be: not exceeding last and not exceeding maximum numbers of SNPs in window set by user
      while(y[i+l] <= win.end && (i+l)<=snps){
        l <- l + 1
      }
      
      # move l back before the threshold
      l <- l - 1
      
      # calculate the number of SNPs in the window
      test <- length(x[i:(i+l)])
      current.window <- (y[l+i] - y[i])
      num.outlier <- sum(z[i:(i+l)] >= top)
      win.mean <- sum(x[i:(i+l)])/(l + 1)
      results[j, ] <- c(test,cont[k],win.start,y[i],win.end,i,(l+1), win.mean, current.window, num.outlier)
      
      # create seperate file with info for every SNP
      for(b in i:(i+l)){
        SNPinfo[a,] <- c(cont[k], y[b], x[b], z[b], win.start, win.end)
        a <- a + 1
      }
      

        win.start <- win.start + slide
        while(y[i] < win.start && i < snps ){
          i <- i + 1
        }
        
        j <- j + 1
        a <- a + 1
        l <- 0
      }
      
      i <- 1
      k <- k + 1 
    }

  results$num.outlier <- as.numeric(results$num.outlier)
  results$current.window <- as.numeric(results$current.window)
  results$win.mean <- as.numeric(results$win.mean)
  results$num.SNPs <- as.numeric(results$num.SNPs)
  hist(results$current.window, breaks = 50)

   write.table(results, file = output) 
   temp <- paste0(output,"_SNPinfo")
   write.table(SNPinfo, file = temp)
}
```

Call function for the different datasets
```{r}

outlier.window(x_na, "xtxna_500-500_top95",.95)
outlier.window(x_eu, "xtxeu_500-500_top95",.95)
outlier.window(x_au, "xtxau_500-500_top95",.95)

outlier.window(x_na, "xtxna_500-500_top99",.99)
outlier.window(x_eu, "xtxeu_500-500_top99",.99)
outlier.window(x_au, "xtxau_500-500_top99",.99)

outlier.window(x_na, "xtxna_50-500_top95",.95)
outlier.window(x_eu, "xtxeu_50-500_top95",.95)
outlier.window(x_au, "xtxau_50-500_top95",.95)

outlier.window(x_na, "xtxna_50-500_top99",.99)
outlier.window(x_eu, "xtxeu_50-500_top99",.99)
outlier.window(x_au, "xtxau_50-500_top99",.99)

outlier.window(x_na, "xtxna_1000_top99",.99)
outlier.window(x_eu, "xtxeu_1000_top99",.99)
outlier.window(x_au, "xtxau_1000_top99",.99)
```

