Pairwise pop Fstats
===

Load in data and read STRUCTURE infiles
```{r, include=FALSE}
library(adegenet) #read.structure & basic stats
library(hierfstat) #pairwise Fst
library(ggplot2) #plotting
library(pegas)
library(car)
library(diveRsity)
library(devtools)
library(plyr) #revalue function (used to rename levels)
library(reshape2)
library(poppr)

allr<-read.structure("allcra473_240_in.stru", n.ind=473, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=85, ask=FALSE, quiet=FALSE)

allrother<-read.table("allr470_other.txt",header=T)
allr@other <- as.data.frame(allrother[,1:6])
allr@pop<-allr@other[,1]
strata(allr)<-data.frame(allr@other[,3:4])
allr@other$latlong<-data.frame(allr@other[,5:6])
```

```{r}
allr_pfst<-pairwise.fst(allr, res.type=c("dist","matrix"))
write.table(as.matrix(allr_pfst),"allr_pfst",sep="\t")
```


Now I wanna know if these pairwise Fsts are significant. So I can permutate across each fst value
From http://lists.r-forge.r-project.org/pipermail/adegenet-forum/2011-February/000214.html

```{r}
mat.obs <- pairwise.fst(allr, res.type="matrix") 
NBPERM <- 10 # this is the number of permutations used for the p-values; for a publication at least 999 would be required.
mat.perm <- lapply(1:NBPERM, function(i) pairwise.fst(allr, pop=sample(pop(allr)), res.type="matrix"))
```

mat.obs contains original Fst values, mat.perm is a list with NPERM matrices of permuted Fst values. To get e.g. right-tail p-values, we can just count the proportion of mat.obs >= mat.perm; e.g. for the first pair of populations:

```{r}
mean(c(mat.obs[1,2] < na.omit(sapply(1:NBPERM, function(i) mat.perm[[i]][1,2])), TRUE))
```

In the above command, "na.omit(sapply(1:NBPERM, function(i) mat.perm[[i]][1,2]))" is a vector of permuted values for this pair of populations across all replicates; c(..., TRUE) is added because the observed value is always added to the permuted values (it is one of the possible permutations of groups).

In practice, it is easier to convert the results as objects of the class randtest (class for Monte Carlo test in ade4, and do for all pairs of populations

```{r}
library(ade4)

allTests <- list()
 for(i in 1:(nrow(mat.obs)-1)){
   for(j in 2:nrow(mat.obs)){
   allTests[[paste(rownames(mat.obs)[i],rownames(mat.obs)[j],sep="-")]] <- as.randtest(na.omit(sapply(1:NBPERM, function(k) mat.perm[[k]][i,j])), mat.obs[i,j], alter="greater")
   }
}
```
