Population descriptions
========================================================

Script by Lotte van Boheemen

Last edit 

```{r}
# Thu Mar 24 13:28:23 2016 ------------------------------
```

Edit 24-3-16

- include PopGenReport

Script to calculate population statistics in R (Fst etc)

Load in data and read STRUCTURE infiles
```{r, include=FALSE}
library(adegenet) #read.structure & basic stats
library(hierfstat) #pairwise Fst
library(ggplot2) #plotting
library(pegas)
library(car)
library(diveRsity)
library(devtools)
library(plyr) #revalue function (used to rename levels)
library(reshape2)
library(poppr)
library(mmod)

allr<-read.structure("allcra473_240_in.stru", n.ind=473, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=85, ask=FALSE, quiet=FALSE)
all<-read.structure("all480_240_in.stru", n.ind=480, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=87, ask=FALSE, quiet=FALSE)
eu<-read.structure("eu199_240_in.stru", n.ind=199, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1, NA.char="-9", pop=36, ask=FALSE, quiet=FALSE)
nar<-read.structure("nacra232_240_in.stru", n.ind=232, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=42, ask=FALSE, quiet=FALSE)
au<-read.structure("au42_240_in.stru", n.ind=42, n.loc=1000, onerowperind=FALSE, col.lab=1, col.pop=2, row.marknames=1,NA.char="-9", pop=7,  ask=FALSE, quiet=FALSE)
```

```{r}
allother<-read.table("all476_other.txt",header=T)
all@other <- as.data.frame(allother[,1:6])
all@pop<-all@other[,1]
strata(all)<-cbind(data.frame(all@other[,1]),data.frame(all@other[,3:4]))
nameStrata(all)<-~Pop/Range/Country
all@other$latlong<-data.frame(all@other[,5:6])

allrother<-read.table("allr470_other.txt",header=T)
allr@other <- as.data.frame(allrother[,1:6])
allr@pop<-allr@other[,1]
strata(allr)<-data.frame(allr@other[,3:4])
allr@other$latlong<-data.frame(allr@other[,5:6])

euother<-read.table("eu197_other.txt",header=T)
eu@other <- as.data.frame(euother[,1:6])
strata(eu)<-data.frame(eu@other[,3:4])
eu@pop<-eu@other[,1]
eu@other$latlong<-data.frame(eu@other[,5:6])

narother<-read.table("nar232_other.txt",header=T)
nar@other <- as.data.frame(narother[,1:6])
strata(nar)<-data.frame(nar@other[,3:4])
nar@pop<-nar@other[,1]
nar@other$latlong<-data.frame(nar@other[,5:6])

auother<-read.table("au41_other.txt",header=T)
au@other <- as.data.frame(auother[,1:6])
strata(au)<-data.frame(au$other[,3:4])
au@pop<-au@other[,1]
au@other$latlong<-data.frame(au@other[,5:6])
```

---
All populations together

````{r}
#au<-read.genetix("au42_1000.gtx")
baseallr<-basic.stats(allr)

```



```{r}
#Gst_Nei(au)

#genind2genpop(au)

#obs.D <- D_Jost(au)
#data("nancycats")

#bs <- chao_bootstrap(nancycats,nreps=10)
#bs.D <- summarise_bootstrap(bs,D_Jost)

#bs <- chao_bootstrap(au,nreps=10) #works
#bs.D <- summarise_bootstrap(bs,D_Jost) #doesn't work
#summarise_bootstrap(chao_bootstrap(au,nreps=10),D_Jost)

Ho_allr<-as.data.frame(baseallr$Ho)
write.table(Ho_allr,file="HoAllr.txt")
Hs_allr<-baseallr$Hs
write.table(Hs_allr,file="HsAllr.txt")
Fis_allr<-baseallr$Fis 
write.table(Fis_allr,file="FisAllr.txt")
n.ind.samp_allr<-baseallr$n.ind.samp
write.table(n.ind.samp_allr,file="nindsamp_allr.txt")
#I used this table to remove loci with less then 4 genotypes
#allelic richness
rich_allr<-allelic.richness(allr,min.n=NULL,diploid=TRUE)
write.table(rich_allr,file="rich_allr.txt")
```

====
**Population level diversity statistics**

Read in data
```{r}
popstat_all<-read.table("popstat_hierfstat.txt", header=TRUE)
popstat_all$N<-as.numeric(popstat_all$N)
heho<-read.table("heho_hierfstat.txt", header=TRUE)
popstat_all$Range <- factor(popstat_all$Range,levels = rev(levels(popstat_all$Range)))
popstat_all$Range<-revalue(popstat_all$Range, c("nat"="Native", "inv_eur"="Introduced, Europe","inv_aus"="Introduced, Australia"))
heho$Range <- factor(heho$Range,levels=rev(levels(heho$Range)))
heho$Range<-revalue(heho$Range, c("nat"="Native", "inv_eur"="Introduced, Europe","inv_aus"="Introduced, Australia"))
heho$Level<- factor(heho$Level, levels=c("Hs", "Ho"))
```

```{r}
pop_nat<-subset(popstat_all, Range=="nat")
shape<-rep(seq(15,18),length.out=42)
pdf("_Graph/LatFis_nat.pdf",height=9,width=9)
ggplot(pop_nat, aes(x=Latitude, y=Fis)) + ylab("Fis")+  
  theme_bw()+theme(legend.position="bottom",legend.direction="horizontal")+
 geom_point(aes(color=factor(PopID)),shape=shape, size=4)+
    geom_errorbar(aes(ymin=Fis-Fis_C, ymax=Fis+Fis_C), width=.2)+
  scale_color_discrete(breaks=c("AA-2","AA-1","NB1","TR-QB","BN-ON","MN-ON","MN-2","AA-3","AA-20","AA-4","MI","AA-19","AA-5","AA-6","WI","MA","AA-17","AA-7","AA-8","AA-16","AA-15","AA-14","AA-9","AA-10","AA-11","AA-12","AA-13","MS","PA","OH","IND","KS","KY","MO","TN","NC","SC","AR","GA","MP","AL","FL"),name="Location code")
dev.off()


pdf("_Graph/LonFis_nat.pdf",height=9,width=9)
ggplot(pop_nat, aes(x=Longitude, y=Fis)) + ylab("Fis")+  
  theme_bw()+theme(legend.position="bottom",legend.direction="horizontal")+
 geom_point(aes(color=factor(PopID)),shape=shape, size=4)+
    geom_errorbar(aes(ymin=Fis-Fis_C, ymax=Fis+Fis_C), width=.2)+
  scale_color_discrete(breaks=c("AA-2","AA-1","NB1","TR-QB","BN-ON","MN-ON","MN-2","AA-3","AA-20","AA-4","MI","AA-19","AA-5","AA-6","WI","MA","AA-17","AA-7","AA-8","AA-16","AA-15","AA-14","AA-9","AA-10","AA-11","AA-12","AA-13","MS","PA","OH","IND","KS","KY","MO","TN","NC","SC","AR","GA","MP","AL","FL"),name="Location code")
dev.off()

pdf("Graph/LatHo_nat.pdf",height=9,width=9)
ggplot(pop_nat, aes(x=Latitude, y=Ho)) + ylab("Observed Heterozygosity")+  
  theme_bw()+ theme(legend.position="bottom",legend.direction="horizontal")+
  geom_point(aes(color=factor(PopID)),shape=shape, size=4)+
    geom_errorbar(aes(ymin=Ho-Ho_C, ymax=Ho+Ho_C), width=.2)+
  scale_color_discrete(breaks=c("AA-2","AA-1","NB1","TR-QB","BN-ON","MN-ON","MN-2","AA-3","AA-20","AA-4","MI","AA-19","AA-5","AA-6","WI","MA","AA-17","AA-7","AA-8","AA-16","AA-15","AA-14","AA-9","AA-10","AA-11","AA-12","AA-13","MS","PA","OH","IND","KS","KY","MO","TN","NC","SC","AR","GA","MP","AL","FL"),name="Location code")
dev.off()

pdf("Graph/LonHo_nat.pdf",height=9,width=9)
ggplot(pop_nat, aes(x=Longitude, y=Ho)) + ylab("Observed Heterozygosity")+  
  theme_bw()+ theme(legend.position="bottom",legend.direction="horizontal")+
  geom_point(aes(color=factor(PopID)),shape=shape, size=4)+
    geom_errorbar(aes(ymin=Ho-Ho_C, ymax=Ho+Ho_C), width=.2)+
  scale_color_discrete(breaks=c("AA-2","AA-1","NB1","TR-QB","BN-ON","MN-ON","MN-2","AA-3","AA-20","AA-4","MI","AA-19","AA-5","AA-6","WI","MA","AA-17","AA-7","AA-8","AA-16","AA-15","AA-14","AA-9","AA-10","AA-11","AA-12","AA-13","MS","PA","OH","IND","KS","KY","MO","TN","NC","SC","AR","GA","MP","AL","FL"),name="Location code")
dev.off()
```


Test for significance between ranges
```{r}
library(lme4)
library(PMCMR)
shapiro.test(popstat_all$Fis)
shapiro.test(popstat_all$AR)
test<-aov(popstat_all$Fis~popstat_all$Range*popstat_all$N)
test1<-aov(popstat_all$AR~popstat_all$Range*popstat_all$N)
summary(test1)
kruskal.test(popstat_all$Fis~popstat_all$Range)
kruskal.test(popstat_all$AR~popstat_all$Range)
posthoc.kruskal.nemenyi.test(popstat_all$AR~popstat_all$Range)
```


Plot results
```{r}
pdf("Graph/HeHo.pdf",height=3,width=3)
ggplot(heho, aes(x=Range, y=H)) + ylab("Heterozygosity")+  
  geom_boxplot(aes(fill=heho$Level))+  theme_bw() +
  theme(legend.justification=c(1,1), legend.position=c(1.1,1.2), legend.text   = element_text( size = 8), axis.title.x = element_text( size=9), axis.text.x  = element_text(size=8,angle=10), axis.title.y = element_text( size=9), axis.text.y  = element_text(size=8)) +scale_fill_discrete(name="",breaks=c("Hs","Ho"),
  labels=c("Expected","Observed"))
dev.off()

pdf("Graph/all_rich.pdf",height=3,width=3)
ggplot(popstat_all, aes(x=Range, y=AR)) + ylab("Allelic richness")+ theme_bw()+ geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))+
theme(axis.title.x = element_text(size=9), axis.text.x  = element_text(size=8,angle=10), axis.title.y = element_text( size=9), axis.text.y  = element_text(size=8))
dev.off()

pdf("_Graph/FIS_all.pdf",height=3,width=3)
ggplot(popstat_all, aes(x=Range, y=Fis)) + ylab("Inbreeding coefficient")+ theme_bw()+ geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))+
theme(axis.title.x = element_text(size=9), axis.text.x  = element_text(size=8,angle=10), axis.title.y = element_text( size=9), axis.text.y  = element_text(size=8))
dev.off()


ggplot(popstat_all, aes(x=Range, y=HE)) + ylab("Expected Heterozygosity")+geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))

ggplot(popstat_all, aes(x=Range, y=HO)) + ylab("Observed Heterozygosity")+geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0))
```


```{r}
popstat_all<-read.table("popstat.txt", header=TRUE)
popstat_h<-read.table("popstat_hierfstat.txt", header=TRUE)

ggplot(popstat_all, aes(x=Range, y=FIS)) + ylab("FIS")+geom_boxplot(outlier.shape=NA) +
  geom_jitter(position=position_jitter(width=.1, height=0))
ggplot(popstat_h, aes(x=Range, y=Fis)) + ylab("FIS")+geom_boxplot(outlier.shape=NA) +
  geom_jitter(position=position_jitter(width=.1, height=0))
ggplot(popstat_all, aes(x=Range, y=FIS)) + ylab("HO")+geom_boxplot(outlier.shape=NA) +
  geom_jitter(position=position_jitter(width=.1, height=0))
ggplot(popstat_h, aes(x=Range, y=Fis)) + ylab("Ho")+geom_boxplot(outlier.shape=NA) +
  geom_jitter(position=position_jitter(width=.1, height=0))
```

===
Compare between ranges (first code ranges as pop). Can't do this according to Kay as I am not accounting for popiulation structure. Therefor, I will get different overall values if I pool all individuals together. 
```{r}
allr@pop<-allr@other[,3]
base_range<-basic.stats(allr)

write.table(base_range$Fis,"loc-range_Fis.txt")

rich_range<-allelic.richness(allr,min.n=NULL,diploid=TRUE)
write.table(rich_allr,file="rich_range.txt")

n.ind.samp_allr<-base$n.ind.samp
write.table(base_range$n.ind.samp,file="nindsamp_range.txt")

```

===

```{r}
library("ape") # To visualize the tree using the "nj" function
library("magrittr")
library("phyclust")
pdf("Graph/try3.pdf", width=20, height=60)
aboot(all, dist = provesti.dist, sample = 2000, tree = "nj", cutoff = 50)
dev.off()

set.seed(999)
all %>% 
  genind2genpop(pop = ~Range/Pop) %>%
 aboot(cutoff = 50, quiet = FALSE, sample = 10, distance=provesti.dist)


all_gp<-genind2genpop(all,pop = ~Range/Pop)
gp_tree<-aboot(all_gp,cutoff = 50, quiet = FALSE, sample = 1000, distance=provesti.dist)
cols <- rainbow(4)
pdf("Graph/unjTree_pops.pdf", width=20, height=60)
plotnj(gp_tree,show.tip.label = TRUE, cex = 1.5, font = 2, lab4ut="axial") #tip.color = cols[all@other$range], 
dev.off()

pdf("Graph/try3.pdf", width=20, height=60)
plot.phylo(gp_tree, cex = 0.8, font = 2, adj = 0,label.offset = 0.0125)#, tip.color = cols[all@other$range]
dev.off()

pdf("Graph/try4.pdf", width=20, height=60)
plotnj(gp_tree,show.tip.label = TRUE, cex = 0.8, font = 2, adj = 0, tip.color = cols[all@other$range],
           label.offset = 0.0125)
dev.off()
```
