Load in data and read STRUCTURE infiles
```{r, include=FALSE}
library(adegenet)
library(hierfstat) #pairwise Fst
library(PMCMR)

popstat_all<-read.table("~/Documents/Monash/PhD/Analyses/Data/popstat_hierfstat.txt", header=TRUE)
```


```{r}
shapiro.test(popstat_all$Fis)
shapiro.test(popstat_all$AR)
```

Fis is non-normal, cannot use anova for this. Will use Kruskal-Wallis (non-parametric test) for all statistics for consistency of methods

Quickly check the effect of sample size on the model. Need to be applicable for non-linear data, but similar to ancova

```{r}
test<-glm(Fis~Range*N, data=popstat_all)
summary(test)
test2<-glm(AR~Range*N, data=popstat_all)
summary(test2)
```
N has no effect on the relation between response and predictors.


Testing:
```{r}
kruskal.test(popstat_all$Fis~popstat_all$Range)
kruskal.test(popstat_all$AR~popstat_all$Range)
```

So allelic richness is different between Ranges, but where is the difference?
```{r}
posthoc.kruskal.nemenyi.test(popstat_all$AR~popstat_all$Range)
```