---
title: "Rmaps"
---

From Kim Gilbert, at:
http://www.molecularecologist.com/2012/09/making-maps-with-r/

Q population values are derived from Structure, StructureHarvester and Clumpp results (K*.popq files)
```{r}
library(maps)
library(mapdata)
library(maptools)  #for shapefiles
library(mapplots)
library(scales)
```



Read in sample points
```{r}
q_all_K2<-read.csv("popq_all_K2.csv")
q_all_K4<-read.csv("popq_all_K4.csv")
q_eu_K4<-read.csv("popq_eu_K4.csv")
q_au_K4<-read.csv("popq_au_K4.csv")
```

Setting up graphics device
```
quartz()
```
Load map (this can better be drawn for every specific part I want to plot)
```{r}
map(database="worldHires",  xlim=c(-110,-60), ylim=c(35,55)) #North America
map(database="worldHires",  xlim=c(-2,30), ylim=c(40,50)) #Europe
```

Here, xlim is longitude, ylim is latitude

For Na and Eu pops together, K=2 (for Europe I use factor 0.02 for N size, 0.07 for Na. )
```{r}
points(q_all_K2$Long,q_all_K2$Lat)
long<-q_all_K2$Long
lats<-q_all_K2$Lat
Q<-q_all_K2$Q
P<-q_all_K2$P
N<-sqrt(q_all_K2$Sample_size*0.02)
lab=q_all_K2$PopID
for(i in 1:87) {
add.pie(z=c(Q[i], P[i]), x=long[i], y=lats[i], radius=N[i], labels=lab[i], cex=0.7,label.dist=2,clockwise=FALSE,col=c(alpha("orange", 0.5), alpha("blue", 0.5)))
}
```

For Na and Eu pops together, K=5
```{r}
points(q_all_K5$Long,q_all_K5$Lat)
long2<-q_all_K5$Long
lats2<-q_all_K5$Lat
Q2<-q_all_K5$Q
P2<-q_all_K5$P
R2<-q_all_K5$R
S2<-q_all_K5$S
T2<-q_all_K5$T
N2<-sqrt(q_all_K5$Sample_size*0.02)
lab2=q_all_K5$PopID
for(i in 1:57) {
add.pie(z=c(Q2[i],P2[i],R2[i],S2[i],T2[i]), x=long2[i], y=lats2[i], radius=N2[i], labels=lab2[i], cex=0.5,label.dist=2,clockwise=FALSE,col=c(alpha("darkgreen",0.5), alpha("orange", 0.5),alpha("yellow",0.5),alpha("red"),0.5,alpha("blue",0.5)))
}
```

This will loop through every row and plot a chart accordingly. The graphical parameters in this code apply to the label call only. The alpha part applies to the transparency. Clockwise depicts how the pies are drawn.

