---
title: "Rmaps"
---

From Kim Gilbert, at:
http://www.molecularecologist.com/2012/09/making-maps-with-r/

Q population values are derived from Structure, StructureHarvester and Clumpp results (K*.popq files)
```{r}
library(maps)
library(mapdata)
library(maptools)  #for shapefiles
library(mapplots)
library(scales)
```



Read in sample points
```{r}
q_all_K2<-read.table("popq_all_K2.txt",header=T)
q_all_K4<-read.table("popq_all_K4.txt",header=T)
q_eu_K4<-read.table("popq_eu_K4.txt",header=T)
q_au_K4<-read.table("popq_au_K4.txt",header=T)
```

Setting up graphics device
```{r}
quartz()
```
Load map (this can better be drawn for every specific part I want to plot)
```{r}
map(database="worldHires",  xlim=c(-110,-60), ylim=c(30,55)) #North America
map(database="worldHires",  xlim=c(-1,30), ylim=c(42,58)) #Europe
map(database="worldHires",  xlim=c(149,156), ylim=c(-38,-25)) #Australia
map(database="worldHires",  xlim=c(-110,156), ylim=c(-38,60)) #world
```

Here, xlim is longitude, ylim is latitude

For all pops (no KSAP or AP1), K=2 (for Europe I use factor 0.02 for N size, 0.07 for Na. )

```{r}
points(q_all_K2$Long,q_all_K2$Lat)
long<-q_all_K2$Long
lats<-q_all_K2$Lat
Q<-q_all_K2$Q
P<-q_all_K2$P
N<-sqrt(q_all_K2$Sample_size*0.02)
lab=q_all_K2$PopID
for(i in 1:85) {
                add.pie(z=c(Q[i], P[i]), x=long[i], y=lats[i], radius=N[i], 
                labels=lab[i], cex=0.7, label.dist=2, clockwise=FALSE, 
                col=c(alpha("orange", 0.5), alpha("blue", 0.5)))
}
```

North America only (k=2)
```{r}
points(q_na_K2$lon,q_na_K2$lat)
long<-q_na_K2$lon
lats<-q_na_K2$lat
Q<-q_na_K2$Q
P<-q_na_K2$P
N<-sqrt(q_na_K2$sample_size*0.02)
lab=q_na_K2$popID
for(i in 1:85) {
add.pie(z=c(Q[i], P[i]), x=long[i], y=lats[i], radius=N[i], labels=lab[i], cex=0.7,label.dist=2,clockwise=FALSE,col=c(alpha("orange", 0.5), alpha("blue", 0.5)))
}
```

Europe only (K=2)
```{r}
points(q_eu_K2$lon,q_eu_K2$lat)
long<-q_eu_K2$lon
lats<-q_eu_K2$lat
Q<-q_eu_K2$Q
P<-q_eu_K2$P
N<-sqrt(q_eu_K2$sample_size*0.02)
lab=q_eu_K2$popID
for(i in 1:85) {
add.pie(z=c(Q[i], P[i]), x=long[i], y=lats[i], radius=N[i], labels=lab[i], cex=0.7,label.dist=2,clockwise=FALSE,col=c(alpha("orange", 0.5), alpha("blue", 0.5)))
}
```

Australia only (K=2)
```{r}
points(q_au_K2$lon,q_au_K2$lat)
long<-q_au_K2$lon
lats<-q_au_K2$lat
Q<-q_au_K2$Q
P<-q_au_K2$P
N<-sqrt(q_au_K2$sample_size*0.02)
lab=q_au_K2$popID
for(i in 1:85) {
add.pie(z=c(Q[i], P[i]), x=long[i], y=lats[i], radius=N[i], labels=lab[i], cex=0.7,label.dist=2,clockwise=FALSE,col=c(alpha("orange", 0.5), alpha("blue", 0.5)))
}
```

For all pops (no KSAP or AP1), K=4
```{r}
points(q_allr_K4$lon,q_allr_K4$lat)
long2<-q_allr_K4$lon
lats2<-q_allr_K4$lat
Q2<-q_allr_K4$Q
P2<-q_allr_K4$P
R2<-q_allr_K4$R
S2<-q_allr_K4$S
N2<-sqrt(q_allr_K4$sample_size*0.02)
lab2=q_allr_K4$popID
for(i in 1:85) {
add.pie(z=c(Q2[i],P2[i],R2[i],S2[i]), x=long2[i], y=lats2[i], radius=N2[i], labels=lab2[i], cex=0.5,label.dist=2,clockwise=FALSE,col=c(alpha("darkgreen",0.5), alpha("orange", 0.5),alpha("yellow",0.5),alpha("red",0.5)))
}
```


This will loop through every row and plot a chart accordingly. The graphical parameters in this code apply to the label call only. The alpha part applies to the transparency. Clockwise depicts how the pies are drawn.
For Na and Eu pops together, K=5

For Europe only (K=4)

```{r}
long2<-q_eu_K4$Long
lats2<-q_eu_K4$Lat
Q2<-q_eu_K4$Q
P2<-q_eu_K4$P
R2<-q_eu_K4$R
S2<-q_eu_K4$S
N2<-sqrt(q_eu_K4$Sample_size*0.02)
lab2<-q_eu_K4$PopID
for(i in 1:36) {
add.pie(z=c(Q2[i],P2[i],R2[i],S2[i]), x=long2[i], y=lats2[i], radius=N2[i], 
        labels=lab2[i],cex=0.5,label.dist=2,clockwise=FALSE,col=c(alpha("darkgreen",0.5), alpha("orange", 0.5),alpha("yellow",0.5),alpha("red",0.5)))
}
```
