Structure analysis
========================================================

```{r, include=FALSE}
library(ggplot2)
```

###############################################################
plotting evanno data after applying structureHarvester.py
High deltaK of 2 might be artificial due to extremely low likelhoods of K = 1. Vigouroux et al 2008

Read in data
```{r}
evanno_all = read.table("all_evanno.txt",header=F,comment.char='#')
evanno_allr = read.table("allcra_evanno.txt",header=F,comment.char='#')
evanno_eu = read.table("eu_evanno.txt",header=F,comment.char='#')
evanno_nar = read.table("nacra_evanno.txt",header=F,comment.char='#')
evanno_au = read.table("au_evanno.txt",header=F,comment.char='#')
evanno_auwes = read.table("auwes_evanno.txt",header=F,comment.char='#')

names(evanno_all) = c("K","reps","mean_LnPK",  "sd_LnPK",  "Ln1K",	"Ln2K",	"Delta_K")
names(evanno_allr) = c("K","reps","mean_LnPK",  "sd_LnPK",  "Ln1K",	"Ln2K",	"Delta_K")
names(evanno_eu) = c("K","reps","mean_LnPK",  "sd_LnPK",  "Ln1K",	"Ln2K",	"Delta_K")
names(evanno_nar) = c("K","reps","mean_LnPK",  "sd_LnPK",  "Ln1K",	"Ln2K",	"Delta_K")
names(evanno_au) = c("K","reps","mean_LnPK",  "sd_LnPK",  "Ln1K",	"Ln2K",	"Delta_K")
names(evanno_auwes) = c("K","reps","mean_LnPK",  "sd_LnPK",  "Ln1K",	"Ln2K",	"Delta_K")
ylab_name<-expression(paste(Delta,"K"))
```

Entire dataset (including KSAP and AP1)
```{r}
all_title<-expression(paste("All samples, including ", italic("A. psilostachya")))
all_Klog<-ggplot(evanno_all, aes(x=K, y=mean_LnPK)) + 
  geom_errorbar(aes(ymin=mean_LnPK-sd_LnPK, ymax=mean_LnPK+sd_LnPK), width=.1) + geom_line() + geom_point() + scale_y_continuous(name="ln Pr(X|K)") + 
  scale_x_continuous(breaks=seq(2, 10,1),name="")+  
ggtitle(all_title) + theme_classic()

all_dK<-ggplot(evanno_all, aes(x=K, y=Delta_K)) + geom_line() + geom_point()+ scale_y_continuous(name=ylab_name)+ scale_x_continuous(breaks=seq(2, 10,1),name="Number of inferred clusters K") + theme_classic()
```


Entire dataset (common ragweed only)
```{r}
titlename<-expression(paste("All ", italic("A. artemisiifolia "), "samples"))
allr_Klog<-ggplot(evanno_allr, aes(x=K, y=mean_LnPK)) + 
  geom_errorbar(aes(ymin=mean_LnPK-sd_LnPK, ymax=mean_LnPK+sd_LnPK), width=.1) + geom_line() + geom_point() + scale_y_continuous(name=" ln Pr(X|K)") + 
  scale_x_continuous(breaks=seq(2, 10,1),name="")+ 
  geom_line() + geom_point() +
  ggtitle(titlename) + theme_classic()

allr_dK<-ggplot(evanno_allr, aes(x=K, y=Delta_K)) + geom_line() + geom_point() +
   scale_y_continuous(name=ylab_name)+ 
  scale_x_continuous(breaks=seq(2, 10,1),name="Number of inferred clusters K") +
  theme_classic()
```


North America (common ragweed only)
```{r}
nar_Klog<- ggplot(evanno_nar, aes(x=K, y=mean_LnPK)) + 
  geom_errorbar(aes(ymin=mean_LnPK-sd_LnPK, ymax=mean_LnPK+sd_LnPK), width=.1) + geom_line() + geom_point() + scale_y_continuous(name="") + 
  scale_x_continuous(breaks=seq(2, 10,1),name="")+    
  geom_line() + geom_point() +
 ggtitle("Native range, North America") + theme_classic()

nar_dK<- ggplot(evanno_nar, aes(x=K, y=Delta_K)) + geom_line() + geom_point() +
  scale_y_continuous(name="")+ 
  scale_x_continuous(breaks=seq(2, 10,1),name="Number of inferred clusters K") + 
  theme_classic()
```

Europe
```{r}
eu_Klog<- ggplot(evanno_eu, aes(x=K, y=mean_LnPK)) + 
  geom_errorbar(aes(ymin=mean_LnPK-sd_LnPK, ymax=mean_LnPK+sd_LnPK), width=.1) + geom_line() + geom_point() + 
  scale_y_continuous(name="") + 
  scale_x_continuous(breaks=seq(2, 10,1),name="")+ 
  ggtitle("Introduced range, Europe") + theme_classic()

eu_dK<- ggplot(evanno_eu, aes(x=K, y=Delta_K)) + geom_line() + geom_point()+ scale_y_continuous(name="")+ scale_x_continuous(breaks=seq(2, 10,1),name="Number of inferred clusters K") +theme_classic()
```

Australia, common ragweed only
```{r}
au_Klog<- ggplot(evanno_au, aes(x=K, y=mean_LnPK)) + 
  geom_errorbar(aes(ymin=mean_LnPK-sd_LnPK, ymax=mean_LnPK+sd_LnPK), width=.1) + geom_line() + geom_point() + scale_y_continuous(name="") +
  scale_x_continuous(breaks=seq(2, 10,1),name="")+  
  ggtitle("Introduced range, Australia") + theme_classic()

au_dK<- ggplot(evanno_au, aes(x=K, y=Delta_K)) + geom_line() + geom_point() + scale_y_continuous(name="")+ scale_x_continuous(breaks=seq(2, 10,1),name="Number of inferred clusters K")  + theme_classic()
```

Australia + western ragweed
```{r}
auwes_Klog<- ggplot(evanno_auwes, aes(x=K, y=mean_LnPK)) + 
  geom_errorbar(aes(ymin=mean_LnPK-sd_LnPK, ymax=mean_LnPK+sd_LnPK), width=.1) + geom_line() + geom_point() + scale_y_continuous(name="") +
  scale_x_continuous(breaks=seq(2, 10,1),name="")+  
  ggtitle("Introduced range, Australia & A. psilostachya") + theme_classic()

auwes_dK<- ggplot(evanno_auwes, aes(x=K, y=Delta_K)) + geom_line() + geom_point() + scale_y_continuous(name="")+ scale_x_continuous(breaks=seq(2, 10,1),name="Number of inferred clusters K")  + theme_classic()
```

Set definitions and plot
```{r}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
pdf("evanno_craplots.pdf", width=20, height=10)
print(multiplot(all_Klog,all_dK,nar_Klog,nar_dK,eu_Klog,eu_dK,au_Klog,au_dK,cols=4))
dev.off()

```
```